---
title: "ReproducibleResearchCourseProject-Satish"
author: "Satish Venkataraman"
date: "September 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)

if(!require(Hmisc)){
  install.packages("Hmisc")
}
library(Hmisc)
```


##1. Code for reading in the dataset and/or processing the data

        Download the file and unzip it to your current working directory.
```{r cache=TRUE}
if(!file.exists("data")){dir.create("data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "./data/repdata.zip")
unzip("./data/repdata.zip")
repData <- read.table("./activity.csv", header = TRUE, sep = ",")
repData$interval<-as.numeric(repData$interval)
repData$steps<-as.numeric(repData$steps)
repData$date<-as.Date(repData$date)
stepsByDay<-aggregate(repData$steps, by=list(repData$date), sum)
colnames(stepsByDay) <- c("dates", "totalSteps")
#stepsByDay$dates<-as.Date(stepsByDay$dates)


```

##2. Histogram of the total number of steps taken each day

```{r}

 hist(stepsByDay$totalSteps,main="Histogram for Total Steps", 
      xlab="day", 
      border="blue", 
      col="green")
#Not sure if they asked for a bar plot since it says number of steps by day, so here goes...
barplot(stepsByDay$totalSteps,names.arg=stepsByDay$dates,main = "Steps By Day", xlab = "Date", ylab = "Total Steps", col = "light green")
```

##3. Mean and median number of steps taken each day
```{r}
statsByDF <- function(x) {
  c(min = min(x), max = max(x), 
    mean = mean(x), median = median(x), 
    std = sd(x))
}
tapply(repData$steps,repData$date,statsByDF)
```

##4. Time series plot of the average number of steps taken
```{r}
timeSeries <- tapply(repData$steps, repData$interval, mean, na.rm = TRUE)

plot(row.names(timeSeries), timeSeries, type = "l", xlab = "5-min interval", 
    ylab = "Average across all Days", main = "Average number of steps", 
    col = "green")
```

##5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
names(which.max(timeSeries))
```

##6. Code to describe and show a strategy for imputing missing data
```{r}
#Making a copy of the data to look at missing values.  Using the mean to impute the missing values

repData.mis<-repData
summary(repData.mis) # This shows 2304 NA values for steps
repData.mis$imputed_steps <- with(repData.mis, impute(steps, mean))
summary(repData.mis) #Checking if the NA values are still shown
```



##7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepsByDayPostImpute<-aggregate(repData.mis$imputed_steps, by=list(repData.mis$date), sum)
colnames(stepsByDayPostImpute) <- c("dates", "totalSteps")
 hist(stepsByDayPostImpute$totalSteps,main="Histogram for Total Steps", 
      xlab="day", 
      border="blue", 
      col="green")
#Not sure if they asked for a bar plot since it says number of steps by day, so here goes...
barplot(stepsByDayPostImpute$totalSteps,names.arg=stepsByDayPostImpute$dates,main = "Steps By Day", xlab = "Date", ylab = "Total Steps", col = "light green")
```
##8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
repData.mis$wDay <- factor((weekdays(repData.mis$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

stepsByDayNew <- aggregate(imputed_steps ~ interval + wDay, data = repData.mis, mean)
names(stepsByDayNew) <- c("interval", "daylevel", "steps")
xyplot(steps ~ interval | daylevel, stepsByDayNew, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```